---
output: github_document
---

[![Travis build status](https://travis-ci.org/USCBiostats/rphyloxml.svg?branch=master)](https://travis-ci.org/USCBiostats/rphyloxml)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# rphyloxml

rphyloxml provides access to the [phyloXML](http://www.phyloxml.org) file format. For now, the only functions that are implemented in the package are:

- `as_phyloxml`: A method to coerce `phylo` objects from the `ape` package as phyloXML (XML) documents.
- `validate_phyloxml`: A wrapper of `xml2::xml_validate`, which allows validating a phyloXML doc using the phyloxml.xsd schema (see [here](http://www.phyloxml.org/1.20/phyloxml.xsd))

This package has been motivated to be used with the javascript library [jsPhyloSVG](http://www.jsphylosvg.org), for which we are currently developing an R package with the same name that provides an htmlwidget [here](https://uscbiostats.github.com/jsPhyloSVG).

## Installation

You can install rphyloxml from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("USCBiostats/rphyloxml")
```

## Example

In the following example, we create a random tree using the rtree function from the ape package, and later on coerce it into a phyloXML document using `as_phyloxml`.

```{r random-tree}
library(ape)
library(rphyloxml)
set.seed(12)

x <- rtree(3)
x
```

```{r coercion}
z <- as_phyloxml(x)
z
```

You can get a "nicer" view of it by doing the following:

```{r nice}
cat(as.character(z))
```

And to store the document, you just need to use `xml2` (which is what powers the package) as follows:

```r
xml2::write_xml(z, "mynicetree.xml")
```




